---
title: "ResearchProject"
output: html_notebook
---

Team members: Vladyslav Humennyy, Volodymyr Kuzma, Svitlana Hovorova.

# The Films Research

We chose a dataset from TMDb, containing various information on films from 1880 to 2022 to explore more about the changes that happened to film industry through the years and test these **4 main hypotheses:**

**1)** The average film duration: we suggest that as time goes films become shorter. H0: on average, the duration of films decreases by 0%; H1: the difference is present. We are also interested in testing this metric in different film categories;

**2)** The frequencies of words used in film titles: we want to find leading words in titles of each year in different genres and see how frequently they occur on the whole timeline;

**3)** The age of leading actors: we assume that in the more recent works, actors are younger. H0: on average, the age of leading actors decreases by 0 years every ten years; H1: the metric is greater;

**4)** The quantities of films of different categories produced: we want to analyze the popularity of genres in particular years.


## Hypothesis 1: Film Duration
```{r}
df = read.csv("TMDb_Dataset.csv")
```

```{r}
Years = seq(1980,2015,length = 8)

Runtimes = c()
for (year in Years) {
  runtime = 0
  num_films = 0
  for (i in 0:4) {
    runtime = runtime + sum(df[grepl(toString(year+i), df$Date), ]$Runtime)
    num_films = num_films + length(df[grepl(toString(year+i), df$Date), ]$Runtime)
    }
  Runtimes = append(Runtimes, runtime/num_films)
}

regressional_model = lm(Runtimes~Years)
plot(Years+2.5, Runtimes, col = "blue")  #
abline(regressional_model, col = "red")
summary(regressional_model)
```
```{r}
Categories = c("Crime", "Comedy", "Action", "Thriller", "Adventure", "Science Fiction", "Drama", "Romance")

for (category in Categories) {
  temp_df = df[grepl(category, df$Categories), ]
  
  Runtimes = c()
  for (year in Years) {
    runtime = 0
    num_films = 0
    for (i in 0:4) {
      runtime = runtime + sum(temp_df[grepl(toString(year+i), temp_df$Date), ]$Runtime)
      num_films = num_films + length(temp_df[grepl(toString(year+i), temp_df$Date), ]$Runtime)
      }
    Runtimes = append(Runtimes, runtime/num_films)
}

  regressional_model = lm(Runtimes~Years)
  plot(Years, Runtimes, col = "blue", main = category)
  abline(regressional_model, col = "red")
  summary(regressional_model)
}
```

